A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-08-17 docs/results_JAMwithNFW_v1.01.tex
started writing section on NFW modelling.

\subsubsection{Sampling with a Markov Monte Carlo Chain}

>>>>

\paragraph{Including a NFW halo.} The dynamical modelling attempts in the previous sections suggest that J1331's inner regions have a slightly more roundish and at large radii more massive mass distribution than expected from the distribution of stars alone. A dark matter halo in addition to the stellar component could explain these findings. We therefore proceed by modelling the mass distribution with a) a stellar component, which we get from the light MGE in Table [TO DO: REF????] times a constant stellar mass-to-light ratio $\Upsilon_*$), and b) a spherical NFW dark matter component [TO DO: Navarro 1995 or 1996???], which is implemented in the JAM modelling as a fit of a 10-Gaussian MGE to the classical NFW profile
\begin{equation*}
\rho_\text{NFW}(r) \propto 1 / \frac{r}{R_\text{s,NFW}} \left( 1 + \frac{r}{R_\text{s,NFW}} \right)^2.
\end{equation*}
The NFW halo has two free parameters, the scale length $R_\text{s,NFW}$ and a parameter describing the total mass of the halo. We use $v_\text{200}$, which is the circular velocity at the radius $r_\text{200}$ within which the mean density of the halo is 200 times the cosmological [TO DO: how to express this] critical density, i.e.
\begin{eqnarray*}
M_\text{200} &=& M(<r_{200})\\
\frac{M_{200}}{ \frac 43 \pi r_{200}^3} &=& 200 \rho_\text{crit}(z=0) \\
v_\text{200} &=& \sqrt{\frac{GM_{200}}{r_\text{200}}}
\end{eqnarray*}
with $\rho_\text{crit}(z=0)=1.43 \cdot 10^{-7} M_\odot / \text{pc}^3$ in the WMAP5 cosmology by [TO DO: REF: Dunkley et al. 2009].

\paragraph{Priors.} [TO DO: REF: Dutton et al. 2010] give a relation for halo vs. stellar mass for late-type galaxies. For an Chabrier IMF stellar population they found
\begin{equation*}
y = y_0 \left( \frac{x}{x_0} \right)^\alpha \left[ \frac 12 + \frac 12 \left(\frac{x}{x_0} \right)^\gamma \right]^{(\beta - \alpha)/\gamma},
\end{equation*}
where $x= m_*$ [$M_\odot h^{-2}$] is the stellar mass, $y = \frac{\langle M_{200}\rangle}{m_*}$ and $\langle M_{200}\rangle$ [$M_\odot h^{-2}$] is the mean halo mass. The parameters for the mean and $2\sigma$ error curves for this relation are$\alpha = -0.5\pm0.15$, $\beta = 0$, $\gamma = 1.0$, $\log_{10} x_0 = 10.4$, $\log_{10} {y_0}_{-0.24}^{+0.28}$ [TO DO: REF: Dutton et al. 2010]. Using the stellar mass estimate for J1331 from \citet{SWELLSI} $m_* = (1.06 \pm 0.25) \cdot 10^{11} M_\odot$ for the Chabrier IMF estimate [TO DO: introduce somewhere], we find ${v_{200}} = (202_{-33}^{+44})_{-13}^{+12}$. The first of the two quoted errors is due to the $2\sigma$ scatter in the relation by [TO DO: REF: Dutton et al. 2010]. The second error is the propagated error due to the uncertainty in the stellar mass. We use this rough estimate for J1331's [TO DO: Continue on page 98]

### 2015-08-17 docs/results_JAMwithSB_v1.01.tex
started writing section on NFW modelling.

\paragraph{... with Increasing Mass-to-Light Ratio.} In \S\ref{sec:results_lensing} we found from the lensing, that the light distribution might drop faster with radius than the mass distribution. This would correspond to a radially increasing total mass-to-light ratio. As velocity anisotropy alone cannot explain the observed kinematics in a simple a mass-follows-light model, we now allow for an mass-to-light ratio gradient in the JAM modelling to generate a mass model from the light distribution in Table \ref{tab:MGEF814W}. We do this by assigning each of the five Gaussians in the MGE its own mass-to-light ratio and replace the total luminosity in Eq. (\ref{eq:centralItotalL}) $L_i$ with the Gaussians total Mass $M_i = \Upsilon_{i} L_i$. We treat the five $\Upsilon_i$ as free parameters and only require that $\Upsilon_{j} \geq \Upsilon_i$ when the corresponding $\sigma_j \geq \sigma_i$ to ensure an overall mass-to-light ratio that is increasing with radius. [TO DO: Discuss Figure]

[TO DO]

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[height=6cm]{fig/jam_G_vrms.ps}
  \caption{Comparison between the stellar $v_\text{rms}$ data (blue points) and the best fit JAM model using an increasing }
  \label{fig:JAM_modelG}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[height=6cm]{fig/jam_G_enclMass.ps}
  \caption{??? [TO DO: Make figure wider. Include legend.???] [TO DO: Why does this plot look different than the others???]}
  \label{fig:???}
\end{subfigure}
\caption{JAM model found by using an increasing mass-to-light ratio to generate a mass model from the light distribution in Table \ref{tab:MGEF814W}. [TO DO]??? Using the surface brightness MGE for the dynamical modelling: each Gaussian got it's own M/L such that the overall M/L is increasing with radius. This is the best fit model. The enclosed mass is overplotted with Einstein mass with 4\% error (overplotted, not fitted). ??? [TO DO: nice caption]}
\label{fig:???}
\end{figure*}

>>>>

\paragraph{... with Increasing Mass-to-Light Ratio.} In \S\ref{sec:results_lensing} we found from the lensing, that the light distribution might drop faster with radius than the mass distribution. This would correspond to a radially increasing total mass-to-light ratio. As velocity anisotropy alone cannot explain the observed kinematics in a simple a mass-follows-light model, we now allow for an mass-to-light ratio gradient in the JAM modelling to generate a mass model from the light distribution in Table \ref{tab:MGEF814W}. We do this by assigning each of the five Gaussians in the MGE its own mass-to-light ratio and replace the total luminosity in Eq. (\ref{eq:centralItotalL}) $L_i$ with the Gaussians total Mass $M_i = \Upsilon_{i} L_i$. We treat the five $\Upsilon_i$ as free parameters and only require that $\Upsilon_{j} \geq \Upsilon_i$ when the corresponding $\sigma_j \geq \sigma_i$ to ensure an overall mass-to-light ratio that is increasing with radius.
\\Figure \ref{fig:enclMass_modelG} shows the (local projected) mass-to-light ratio gradient generated by the best fit to the dynamics data, which rises from $\Upsilon_\text{tot} = 2.53$ in the center and approaches a value of $\Upsilon_\text{tot}$ outside of the fitted region at $R'\gtrsim 3$ arcsec of $\Upsilon_\text{tot} = 7.60)$. The central mass-to-light ratio corresponds to the Chabrier IMF stellar mass-to-light ratio found by [TO DO: REF: TREU ET AL 2011] $\Upsilon_{i,*}^\text{Chab} = 2.5 \pm 0.6$ [TO DO: introduce and explain????]. The strong rise  of$\Upsilon_\text{tot}(R')$ cannot be explained by an increase in the stellar $\Upsilon_*$ alone [TO DO: Really??? Maybe it would have dropped again in the blue disk, if I had allowed for it????], so we deduce that we might need  contribution of dark matter halo in J1331.
\\Fig. \ref{fig:JAM_modelG} shows that the best fit model greatly reproduces the central dip in the $v_\text{rms}$ curve, even though it has slight problems fitting the drop around $R' \sim 4$ arcsec. The latter might be because we only allowed the $\Upsilon_\text{tot}(R')$ to rise. A slight drop could be expected when the reddish bulge turns into the bluish disk and the contribution of the stellar component becomes less due to a lower $\Upsilon_{*}$ for younger and bluer populations.
\\In Fig. \ref{fig:enclMass_modelG} we overplot the enclosed mass profile with the Einstein mass $M_\text{ein} = (7.77 \pm 0.33) \cdot 10^{10} M_\odot$ at the Einstein radius found from lensing in Table [TO DO: REF]. The agreement between the Einstein mass and the independently found $M(<R_\text{ein}) = 7.49 \cdot 10^{10} M_\odot$ from dynamical modelling is striking.

@@

\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[height=6cm]{fig/jam_G_vrms.ps}
  \caption{Comparison of $v_\text{rms}$ data and best fit model.}
  \label{fig:JAM_modelG}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[height=6cm]{fig/jam_G_enclMass.ps}
  \caption{Enclosed mass and mass-to-light ratio profile. [TO DO: Make figure wider. Include legend.???] [TO DO: weird  ticks on right axis???] [TO DO: Why does this plot look different than the others???]}
  \label{fig:enclMass_modelG}
\end{subfigure}
\caption{JAM model found by fitting an increasing total mass-to-light ratio $\Upsilon_\text{tot}$ profile used to generate a mass model from the light distribution.  This is done by assigning a different mass-to-light ratio to each Gaussian in the MGE in Table \ref{tab:MGEF814W}. \emph{Panel (a):} Comparison between the stellar $v_\text{rms}$ data (blue points) and the best fit model (green line). \emph{Panel (b):} Enclosed mass inside the projected radius $R'$ on the sky (green line, left axis) and mass-to-light profile $\Upsilon(R')$ along the major axis [TO DO: CHECK???] (blue line, right axis) of the best fit model. The enclosed mass curve is overplotted with the independent finding for the Einstein mass $\pm 4 \%$ in Table [TO DO: TABLE REFERENCE???] (red dot) at the Einstein radius (red dotted line). The best fit mass-to-light ratios of the first four Gaussians are plotted against each Gaussians $\sigma$ (yellow stars). The two Gaussians with the largest $\sigma$ (the fifth is not shown) have the same best fit mass-to-light ratio. Overplotted is also the effective half-light ratio $R_\text{eff}$ (black dotted line).}
\label{fig:modelG}
\end{figure*}
